Stochastic Approximation Monte Carlo with a Dynamic Update Factor (SAD)

### 2019-03-19 papers/histogram/sad.tex
a little more sad writing

Wang and Landau introduced an algorithm (WL) that uses an update
factor and a statistical histogram to compute the density of states of
a given system~\cite{wang2001determining, wang2001efficient}.  While
the method is incredibly powerful, it has a few disadvantage.
Firstly, it requires the user to select a number of parameters,
including the range of energies to be studied~\cite{landau2004new}.
This adds an additional hurdle to its application to a variety of
systems.  The simulation violates detailed balance, with the size of
the violation decreasing with time, which complicates convergence
analysis.  In fact, the error in a WL computation the error has been
demonstrated to saturate at a non-zero value~\cite{yan2003fast}, i.e. the method does not
converge to the true density of states~\cite{belardinelli2007wang,
belardinelli2008analysis, belardinelli2014intrinsic,
singh2012density}.

Around the same time, the convergence rate for WL was being
explored~\cite{zhou2005understanding,lee2006convergence,
belardinelli2007wang}. It was found that utilizing modification factors
that decrease faster than $1/t^2$ leads to
nonconvergence~\cite{belardinelli2007fast}.  This leads to the so-called
$1/t$-WL algorithm which ensures that excess CPU time is not wasted by
continuing to perform calculations once the error in the density of
states becomes saturated~\cite{belardinelli2008analysis}. An effort to
reduce the number of user defined parameters of WL and formulate a
theory for why the method converges despite detailed balance being
violated although infrequently was also being undertaken.

>>>>

Wang and Landau introduced an algorithm (WL) that uses an update
factor and a statistical histogram to compute the density of states of
a given system~\cite{wang2001determining, wang2001efficient}.  While
the method is incredibly powerful, it has a few disadvantage.
Firstly, it requires the user to select a number of parameters,
including the range of energies to be studied~\cite{landau2004new}.
This adds an additional hurdle to its application to a variety of
systems.  The simulation violates detailed balance, with the size of
the violation decreasing with time, which complicates convergence
analysis.  In fact, the error in a WL computation the error has been
demonstrated to saturate at a non-zero value~\cite{yan2003fast},
i.e. the method does not converge to the true density of
states~\cite{belardinelli2007wang, belardinelli2007fast,
  belardinelli2008analysis, belardinelli2014intrinsic,
  singh2012density}.  
%
Belardinelli and Pereyra demonstrated that allowing the update factor
to decrease faster than $1/t$ leads to
nonconvergence~\cite{belardinelli2007wang}.
%% Around the same time, the convergence rate for WL was being
%% explored~\cite{zhou2005understanding,lee2006convergence,
%%   belardinelli2007wang}.
%
This leads to their $1/t$-WL algorithm which ensures that the error
continues to decrease asymptotically as
$1/\sqrt{t}$~\cite{belardinelli2008analysis}. \davidsays{I don't
  understand the following: An effort to reduce the number of user
  defined parameters of WL and formulate a theory for why the method
  converges despite detailed balance being violated although
  infrequently was also being undertaken.  CUT THIS MAYBE?}
